Solving the Poisson-equation in one dimension \\ \hrulefill\small{ FYS3150: Computational Physics }\hrulefill

### 2016-09-15 Project1/article/article.tex
Started results section

The previous equation can be solved in two steps
\begin{equation}
L\vb{w}=\vb{f}, \quad U\vb{v}=\vb{w}
\end{equation}

@@

@@

\begin{thebibliography}{10}
  \bibitem{thomasalgo}{Thomas, L.H. (1949), \emph{Elliptic
        Problems in Linear Differential Equations over a
        Network}. Watson Sci. Comput. Lab Report, Columbia University,
      New York.}
    \bibitem{morten}{Hjorth-Jensen, M. (2016). \emph{Computational
        Physics - Lecture Notes 2016}. University of Oslo}
\end{thebibliography}

>>>>

The equation $A\vb{v}=LU\vb{v}=\vb{f}$ can be solved in two steps
\begin{equation}
L\vb{w}=\vb{f}, \quad U\vb{v}=\vb{w}
\end{equation}
Written out, the full set of these linear equations will look like
\begin{align}
w_1 &= f_1 \label{eq:LU1}\\
l_{21}w_1 + w_2 &= f_2 \\
l_{31}w_1 + l_{32}w_2 +w_3 &= f_3 \\
 & \vdots \\
l_{n-11}w_1 + l_{n-12}w_2 + l_{n-13}w_3 + \dots + w_{n-1}  &=
                                                                    f_{n-1}\\
l_{n1}w_1 + l_{n2}w_2 + l_{n3}w_3 + \dots + l_{nn-1}w_{n-1} + w_n &=
                                                                    f_n
  \label{eq:LU2}\\  
u_{11}v_1 + u_{12}v_2 + u_{13}v_3 + \dots + u_{1n-1}v_{n-1} +
  u_{1n}v_n &= w_1  \label{eq:LU4}\\  
u_{22}v_2 + u_{23}v_3 + \dots + u_{2n-1}v_{n-1} + u_{2n}v_n &= w_2 \\
u_{33}v_3 + \dots+u_{3n-1}v_{n-1} + u_{3n}v_n &= w_3 \\
 & \vdots \\
u_{n-1n-1}v_{n-1} &=  w_{n-1} \\
u_{nn}v_n &= w_n \label{eq:LU3}
\end{align}
First $\vb{w}$ can be determined by iterating forwards through
equations \ref{eq:LU1} to \ref{eq:LU2}. Then $\vb{v}$ can be found be
iterating backwards from equation \ref{eq:LU3} to equation
\ref{eq:LU4}.

In this study the Armadillo library
is employed instead of implementing the LU decomposition
algorithm from scratch. The LU method is used as a benchmark against which the
speed of the tridiagonal method is measured. The number of FLOPS
required by the LU is $2n^3/3$, by any alogorithm used to implement
it\cite{golub}. Additionally, $n^2$ FLOPS is required by the forward
and backwards substitution.
\begin{equation}
N_{LU}=\frac{2}{3}n^3-2n^2=\mathcal{O}(\frac{2}{3}n^3)
\end{equation}
Using LU computation results in a quadratic slowdown compared to the
tridiagonal matrix algorithm.

@@

\begin{table}[h]
\caption{Elapsed time for different $n$}
\begin{tabular}{lcc}
\hline
n & TDMA [s] & LU [s] \\ \hline
10 & 0.0000035 & 0.00312535 \\
100 & 0.0000116 & 0.00193754 \\
1000 & 0.000077892 & 0.0677764 \\
10000 & 0.000878769 & 21.9247 \\
100000 & 0.00757418 & n/a \\
1000000 & 0.08616075 & n/a \\
10000000 & 0.76534 & n/a
\end{tabular}
\label{tab:solver_times}
\end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/10.png}
  \caption{Numerical approximation for $n=10$}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/100.png}
  \caption{Numerical approximation for $n=100$}
\end{figure}

@@

\begin{thebibliography}{10}
  \bibitem{thomasalgo}{Thomas, L.H. (1949), \emph{Elliptic
        Problems in Linear Differential Equations over a
        Network}. Watson Sci. Comput. Lab Report, Columbia University,
      New York.}
    \bibitem{morten}{Hjorth-Jensen, M. (2016). \emph{Computational
        Physics - Lecture Notes 2016}. University of Oslo}
    \bibitem{golub}{Golub, G.H., van Loan, C.F. (1996). \emph{Matrix
          Computations} (3rd ed.), Baltimore: John Hopkins.}
\end{thebibliography}
